---
title: "Week 04 Assignment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Exercises

1. Use the following code chunk to get started with the assignment. This chunk loads the `tidyverse` and `here` packages. Why does this work? What is the `pacman` package and why is it valuable?

```{r}
if(! require("pacman")) install.packages("pacman")

pacman::p_load(
  tidyverse,
  here
)
```
The provided code works in 2 steps:
1. Check for `pacman` and load or install as necessary
The require function loads an existing package or returns False if the package is missing. For the if statement to proceed it must have a condition met, or logically True. In this case, we want the `install` behavior when `require` is not-True, so we negate the result of `require` with a leading `!` or NOT.

2. Use the p_load command from the pacman package to load the listed libraries to the active session.

`pacman` is a package management suite for R. It provides a wide array of functions that can be used to install, uninstall, load, unload, or retrieve information on packages using short commands.


2. Read the `burritos.csv` file located in the `Data` folder into an object called `burritos`.

```{r}
burritos <- read_csv(here::here("Data", "burritos.csv"),
                     col_types = cols(Queso = "c") 
                     # correction to incorrect default col guess(es)
                     )
glimpse(burritos)
```

3. Determine how many `NA` values exist in each column.

```{r}
burritos %>%
    summarize(across(everything(),~sum(is.na(.))))

```

4. Many of the columns in the dataset use `x` to indicate whether or not the burrito includes the ingredient or not. Convert these columns into logical columns. Overwrite the original dataframe with the newly mutated dataframe.

```{r}
burritos <-
burritos %>%
  mutate(across(c(Chips, Unreliable:Zucchini),
                  ~if_else(is.na(.), F, T)
                )
         ) # %>%
#  summary()

```

5. For these same fields, determine how many burritos include each ingredient.

```{r}

burritos %>%
  select(c(Chips, Beef:Zucchini))  %>%   # anything edible is an ingredient
  summarize(colnames(.),colSums(.))
  # Sum of all logical values = number of entries with value "True"

```

6. Several of the burritos are missing Yelp and Google review values. Replace the missing values by imputing them with the *median* of their respective column. Overwrite the original dataframe with the newly mutated dataframe.

```{r}
burritos <-
burritos %>%
  mutate(Yelp =  
             if_else(is.na(Yelp),     # call function
                     true =  median(Yelp, na.rm = T), 
                     false = Yelp)
         )  %>%
  mutate(Google =  
             if_else(is.na(Google),     # call function
                     true =  median(Google, na.rm = T), 
                     false = Google)
         ) # %>%
  # select(Yelp, Google)

```

7. Similarly, several of the burritos are missing `Length`, `Circum`, and `Volume` values. Replace the missing values by imputing them with the *mean* of their respective column. Overwrite the original dataframe with the newly mutated dataframe.

```{r}
burritos <-
burritos %>%
  mutate(across(c(Length, Circum, Volume),   
             ~if_else(is.na(.x),     # use .x to specify column rather than overall for all columns
                     true =  mean(.x, na.rm = T), 
                     false = .x)
                )
         ) # %>%
#  select(Length, Circum, Volume)
```
At 3 column items, repeating is a bit tedious. There has to be a way to repeat the operation within the column for multiple columns, and so there is -- using ".x" instead of ".".

Noted that first 75 rows are NA in these 3 columns. Check data for correct operation (interpolated vs unaltered) below that line.


8. The `Rec` column is messy. Convert it to a logical field. Overwrite the original dataframe with the newly mutated dataframe.
```{r}
unique(burritos$Rec)  # identify values in column
```

```{r}
#rewrite data from permutations of yes/no/blank_NA to T/F/NA (logical)
burritos <-
burritos %>%
  mutate(Rec =  case_when(
                  Rec == "yes" ~ TRUE, # no quotes for logical value
                  Rec == "Yes" ~ TRUE, 
                  Rec == "Y" ~ TRUE,
                  Rec == "no" ~ FALSE,
                  Rec == "No" ~ FALSE,
                  Rec == "N" ~ FALSE
                                      # NA was OK
                  )
             
         ) #  %>%
  # select(Rec)
```

9. Create and answer a question to answer about the burritos dataset utilizing any of the techniques we discussed in class this week. Utilize themes to improve your visualization.

Q. Do ratings differ between burritos that get a Y/N recommendation?
```{r}
burritos %>% 
ggplot(aes(x = overall)) +
  geom_histogram(na.rm = T,    # Remove warning re 2 missing values
                 bins = 25) +  # 25 rating divisions of 0.2 units
  facet_grid(Rec ~ .) +
  theme_minimal() +
  labs(x = "Overall Rating",
       y = "Count",
       title = "Rating Distribution by Recommendation Status")
```
Overall, burritos that get recommended have higher ratings than ones that do not. The  burritos with a rating but not a recommendation status (NA) seem to cover the full range of possible ratings.

Let's get numerical to confirm...
```{r}
burritos %>% 
  select(Rec, overall) %>%
  group_by(Rec) %>%
  summarise(count = n(), mean = mean(overall, na.rm = T), median = median(overall, na.rm = T), stdev = sd(overall, na.rm = T)) %>%
  ungroup()

```
The numerical summary indicates that
a) the mean and median values are very near each other for all 3 categories for recommendation (T, F, NA), and 
b) NA has a broader spread than T and F based on standard deviation.
These are consistent with the distribution plots.

For ease of comparison, we'll look at median as the center of the distribution. A burrito that was recommended had a median rating 1.3 points (of 5) greater than one that was not-recommended. 
```{r}
# what is weighted overall rating of burritos with T/F rating? (manual -- could probably program selecting non-NA subset of Rec?)
((70 * 2.7) + (163 * 4.0))/(163+70)
```
The burritos without recommendation data have a median slightly below the ones with positive recommendation. Comparing to the population-weighted median of values with recommendation status, the median without recommendation data is comparable (3.6 predicted vs 3.8 actual) within the spread of the data.

